---
title: "Stat 6021: Project 1"
author:
  - "Gaurav Anand"
  - "Haizhu Hong"
  - "Emily Murphy"
  - "Annie Williams"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
data<-read.csv(file='clean_diamond_data.csv')
attach(data)
```

# Executive Summary




# Introduction to the Data and Our Question

<!-- The audience for the detailed description is another classmate your client may hire to -->
<!-- give a second opinion for your report. -->

<!-- ### I. Summary of continuous and categorical predictors  -->

<!-- ### II. Number of values for each categorical predictor -->

<!-- ### III. Goals -->


The goal of this report is to understand which predictors out of carat, clarity, color, and cut are the most useful in predicting the price of a diamond, as well as understand some of the relationships between these predictors.

The dataset contains 210,638 different diamonds that are for sale on [bluenile.com](https://www.bluenile.com/)

describe categorical variables based on following analysis that we decide on (do we combine categories??)

# Exploratory Analysis

### I. Distributions of values 

### II. Scatterplots (Predictors)

### III. Scatterplots (Price vs. Carat)



# Detailed Analysis

### I. Fitting the Full Model

Try fit full model without transformation

```{r no_transformations, echo=FALSE}
fullreg <- lm(price~carat+clarity+color+cut)
summary(fullreg)
anova(fullreg)
par(mfrow=c(1,3))
plot(fullreg$fitted.values,fullreg$residuals)
abline(h=0,col="red")
acf(fullreg$residuals)
# all visuals in one chunk or on separate pages? how much room do we want to take up? 
```

```{r no_transformations_qq, echo=FALSE}
qqnorm(fullreg$residuals)
qqline(fullreg$residuals,col="red")
boxplot(price~clarity)
```

Try regress price against carat and clarity

```{r echo=FALSE}
clarityreg <- lm(price~carat+clarity)
summary(clarityreg)
anova(clarityreg)
par(mfrow=c(1,2))
plot(clarityreg$fitted.values,clarityreg$residuals)
abline(h=0,col="red")
acf(clarityreg$residuals)
qqnorm(clarityreg$residuals)
qqline(clarityreg$residuals,col="red")
boxplot(price~clarity)
library(lawstat) 
levene.test(price,clarity)
```

### II. VIF Analysis

### III. Residual plot, ACF, QQ Plots

### IV. Potential Transformations



# Conclusions
